---
title: 키움 API Tick Data Cleaning (1)
author: 'Roh'
date: '2021-01-12'
slug: 키움-api-tick-data-cleaning
categories:
  - finance
  - R
  - highfrequency
tags: []
output:
  blogdown::html_page:
    toc: true
    toc_depth: 2
---
<style>
div.rcode pre { background-color:#fffff0; }
div.rcode pre.r { background-color:#EDF6FA; }
</style>

<style>
div.pycode pre { background-color:#fffff0; }
div.pycode pre.python { background-color:#FFE9E9; }
</style>

**이 페이지는 키움 API를 통해 Real Time Tick Data를 받아온 후 데이터를 클리닝 하는 과정을 보여주고 있습니다. 오류 및 문의사항은 metrics@kakao.com 으로 메일주시면 감사하겠습니다**

**데이터 자체에 대한 질문과 데이터 제공에 관한 문의는 000 으로 메일 주시면 감사하겠습니다**

**R code 블럭과 Python code 블럭은 다음과 같이 색깔로 구분하겠습니다. 결과창은 동일하게 Ivory 색 블럭으로 표시됩니다.**

<div class = "rcode">
```{r}
# "이것은 R 코드 입니다."
```
</div>

<div class = "pycode">
```{python}
# "이것은 Python 코드 입니다."
```
</div>

### Package
```{r, warning=FALSE, message=FALSE, error=FALSE}
library(dplyr)
library(DT)
library(reticulate) # Python
#py_install(packages = "matplotlib")
#py_install(packages = "pandas")
#py_install(packages = 'dfply')

options(scipen=999)
options(max.print = 99999999)
options(digits=10)
```

## Import Data
<div class = "rcode">
```{r, warning=FALSE, message=FALSE, error=FALSE}
wd = "G:/공유 드라이브/Project_TBD/Stock_Data/real_time/kiwoom_stocks/2021-01-12"
fn = list.files(path = wd,
                pattern = '.*stocks_trade.*.*_15.*\\.csv') # 마지막 30분 틱데이터 가지고오기
path = paste(wd,fn,sep = '/')
data = readr::read_csv(file = path, 
                       col_names = c('code','trade_date','timestamp','price','open','high','low',
                                     'size','cum_size','ask1','bid1'))
sum(is.na(data))
```
</div>

<div class = "pycode">
```{python}
import numpy as np
import matplotlib.pyplot as plt
import pandas as pd
from datetime import datetime, timedelta
#py_install(packages = "matplotlib")
#py_install(packages = "pandas")
pd.options.display.float_format = '{:.4f}'.format
df_py = r.data
df_py.head()
```
</div>

## Filter for "Samsun Electronics" Stocks

- Code = '005930'

<div class = "pycode">
```{python}
ss = df_py[df_py.code == '005930'].reset_index(drop=True)
ss.head()
```
</div>

## Time Data Cleaning

Tasks to be done

1. Transform data type to str to use substring to make the format like "%Y-%m-%d %H:%M:%S.%f"
2. Delete where second does not line between [0:59]
3. Delete where time difference goes beyond one second between Kiwoom api time and self-recorded time 

<div class = "pycode">
```{python}
# Task 1 & 2
ss = ss[ss['timestamp'].apply(lambda x: int(str(int(x*1000000))[12:14])) < 60 ]

# Task 3
diff = ss['timestamp'].apply(lambda x: int(str(int(x*1000000))[8:14])) - ss['trade_date']
print(diff.value_counts())
ss = ss[abs(diff) <=1]

ss = ss.assign(time = ss['timestamp'].apply(lambda x: datetime(year = int(str(int(x*1000000))[0:4]),
                month = int(str(int(x*1000000))[4:6]),
                day = int(str(int(x*1000000))[6:8]),
                hour = int(str(int(x*1000000))[8:10]),
                minute = int(str(int(x*1000000))[10:12]),
                second = int(str(int(x*1000000))[12:14]),
                microsecond = int(str(int(x*1000000))[14:20])).strftime("%H:%M:%S.%f")[:-3]
               )

)
```
</div>

## Get Trade Direction Indicator

If price is large than midprice, then it is buyer-initiated.
If price is less  than midprice, then it is seller-initated.
It is tricky if price is same as midprice. In that case, we use the following rule.

> The tick rule is the most commonly used level-1 algorithm. This rule is rather simple and classifies a trade as buyer-initiated if the trade price is above the preceding trade price (an uptick trade) and as seller-initiated if the trade price is below the preceding trade price (a downtick trade). If the trade price is the same as the previous trade price (a zero-tick trade), the rule looks for the closest prior price that differs from the current trade price. Zero-uptick trades are classified as buys, and zero-downtick trades are classified as sells.

The rule above can be simply implemented through the code below.

<div class = "pycode">
```{python}
ss = ss.assign(spread = ss['ask1'] - ss['bid1'],
               mid = (ss['bid1'] + ss['ask1'])*.5,
               ind = 0,
               price1 = ss['price'].shift(1, fill_value=0),
               price2 = ss['price'].shift(2, fill_value=0)
              )
              
buy = ((ss['price'] > ss['mid']) |
        ((ss['price'] == ss['mid']) & (ss['price'] > ss['price1'])) |
        ((ss['price'] == ss['mid']) & (ss['price'] == ss['price1']) & (ss['price'] > ss['price2'])))
        
ss.loc[buy, 'ind'] = 1
ss.loc[~buy, 'ind'] = -1

ss1 = ss[['time', 'price','size','ask1','bid1','ind','spread']]

ss1.head(5)
```
</div>


### We will discuss what to do more with using trading indicator value along with other variables extracted from tick data.

## Source

- 데이터 provided by 00 Team
- http://dee.uib.es/digitalAssets/234/234006_Pascual.pdf