---
title: 키움 API Tick Data - Day2
author: Roh
date: '2021-01-14'
slug: api-tick-data-day2
categories:
  - finance
  - Python
  - highfrequency
tags: []
output:
  blogdown::html_page:
    toc: true
    toc_depth: 2
---
<style>
div.rcode pre { background-color:#fffff0; }
div.rcode pre.r { background-color:#EDF6FA; }
</style>

<style>
div.pycode pre { background-color:#fffff0; }
div.pycode pre.python { background-color:#FFE9E9; }
</style>

<style type="text/css">

body, td {
   font-size: 20px;
}
code.r{
  font-size: 12px;
}
pre {
  font-size: 18px
}
</style>

*__Note__: 이 페이지는 키움 API를 통해 Real Time Tick Data를 받아온 후 데이터를 클리닝 하는 과정을 보여주고 있습니다. 오류 및 문의사항은 metrics@kakao.com 으로 메일주시면 감사하겠습니다*

**데이터 자체에 대한 질문과 데이터 제공에 관한 문의는 000 으로 메일 주시면 감사하겠습니다**

**R code 블럭과 Python code 블럭은 다음과 같이 색깔로 구분하겠습니다. 결과창은 동일하게 Ivory 색 블럭으로 표시됩니다.**

<div class = "rcode">
```{r}
# "이것은 R 코드 입니다."
```
</div>

<div class = "pycode">
```{python}
# "이것은 Python 코드 입니다."
```
</div>

### Package
```{r, warning=FALSE, message=FALSE, error=FALSE}
library(dplyr)
library(tidyverse)
library(DT)
library(reticulate) # Python
#py_install(packages = "matplotlib")
#py_install(packages = "pandas")
#py_install(packages = 'dfply')

options(scipen=999)
options(max.print = 99999999)
options(digits=10)
```

### Import Data

- Check NA

<div class = "rcode">
```{r, warning=FALSE, message=FALSE, error=FALSE}
wd = "G:/공유 드라이브/Project_TBD/Stock_Data/real_time/kiwoom_stocks/2021-01-14"
wd2 = "/Volumes/GoogleDrive/공유 드라이브/Project_TBD/Stock_Data/real_time/kiwoom_stocks/2021-01-14"
fn = list.files(path = wd2,
                pattern = '.*stocks_trade.*\\.csv') # 마지막 30분 틱데이터 가지고오기
path = paste(wd2,fn,sep = '/')

# data = readr::read_csv(file = path, 
#                        col_names = c('code','trade_date','timestamp','price','open','high','low',
#                                      'size','cum_size','ask1','bid1'))

tbl =
  list.files(path = wd2, pattern = '.*stocks_trade.*\\.csv') %>%
  map_df(~readr::read_csv(paste(wd2,.,sep = '/'),
                   col_names = c('code','trade_date','timestamp','price','open','high','low',
                                     'size','cum_size','ask1','bid1') ))

sum(is.na(tbl))
```
</div>

<div class = "pycode">
```{python}
import numpy as np
import matplotlib.pyplot as plt
import pandas as pd
from datetime import datetime, timedelta
#py_install(packages = "matplotlib")
#py_install(packages = "pandas")
pd.options.display.float_format = '{:.4f}'.format
pd.set_option('display.max_rows', 100)
df_py = r.tbl
df_py.head()
```
</div>

```{r, include=FALSE, warning=FALSE, message=FALSE, error=FALSE}
rm(list = "tbl")
```

# Contents

## 1. Filter for a certain Stocks [특정주식으로 필터링 하기]

- KOSPI Code for Samsung Electronics is '005930'

<div class = "pycode">
```{python}
ss = df_py[df_py.code == '005930'].reset_index(drop=True)
ss.shape
```
</div>

## 2. Use time object in python

* Delete rows where "second" does not lie between 0 and 59.

* Substring the timstamp to get each time component.

<div class = "pycode">
```{python}

ss = ss[ss['timestamp'].apply(lambda x: int(str(int(x*1000000))[12:14])) < 60 ]

ss = ss.assign( microsecond = ss['timestamp'].apply(lambda x: str(int(x*1000000))[14:20]),
                second = ss['timestamp'].apply(lambda x: str(int(x*1000000))[12:14]),
                minute = ss['timestamp'].apply(lambda x: str(int(x*1000000))[10:12]),
                hour = ss['timestamp'].apply(lambda x: str(int(x*1000000))[8:10]),
                day = ss['timestamp'].apply(lambda x: str(int(x*1000000))[6:8]),
                month = ss['timestamp'].apply(lambda x: str(int(x*1000000))[4:6]),
                year = ss['timestamp'].apply(lambda x: str(int(x*1000000))[0:4]),
                kw_time = ss['trade_date'].apply(lambda x: str(int(x))))
ss.head()
```
</div>

* Transform to python datetime!

<div class = "pycode">
```{python}
ss['time'] = ss.apply(lambda x: datetime(
                  year = int(x['year']),
                  month = int(x['month']),
                  day = int(x['day']),
                  hour = int(x['hour']),
                  minute = int(x['minute']),
                  second = int(x['second']),
                  microsecond = int(x['microsecond'])), #.strftime("%H:%M:%S.%f")[:-3], 
                  axis=1)
ss.head(10)
```
</div>

* Calculate the latency

<div class = "pycode">
```{python}
ss['kwtime'] = ss.apply(lambda x: datetime(year = int(x['year']),
                                           month = int(x['month']),
                                           day = int(x['day']),
                                           hour = int(x['kw_time'][0:2]),
                                           minute = int(x['kw_time'][2:4]),
                                           second = int(x['kw_time'][4:6])),
                                           axis = 1
                                           )
ss['latency'] = ss.apply(lambda x: x['time'].second - x['kwtime'].second,
                          axis =1
                          )
ss1 = ss[['time', 'kwtime', 'hour', 'minute', 'second', 'price','size','ask1','bid1','latency']]
ss1.head(10)
```
</div>


## 3. Calculate latency and clean it [레이턴시 계산 및 정리]

*Calculate the latency by substracting the tiem provided by Kiwoom dataset and self-recorded time at the moment of processing the data delivered through the Kiwoom API.

<div class = "pycode">
```{python}
# Task 2 & 3
print(ss1.latency.value_counts())
print(ss1.latency.nsmallest(10))
print(ss1.latency.nlargest(10))
```
</div>

<div class = "pycode">
```{python}
plt.plot(ss1.latency)
plt.title('Latency Over Time Before Cleaning')
```
</div>

* Check the time where large discrepancy happened!

#### When latency is -59/59, it actually is 1 second difference

<div class = "pycode">
```{python}
ss1[ss1.latency == -59][['kwtime','hour','minute','second','latency']]
```
</div>

#### When latency is 59, it actually is 1 second difference

<div class = "pycode">
```{python}
ss1[ss1.latency == 59][['kwtime','hour','minute','second','latency']]
```
</div>

* Delete where time difference goes beyond one second between Kiwoom api time and self-recorded time. In other words, we allow latency +/- five second at maximum.

<div class = "pycode">
```{python}
ss2 = ss1[abs(ss1.latency) <=5]
print(ss2.latency.value_counts())

plt.plot(ss2.latency)
plt.title('Latency Over Time After Cleaning')
```
</div>

## 4. Get Trade Direction Indicator

If price is large than midprice, then it is buyer-initiated.
If price is less  than midprice, then it is seller-initated.
It is tricky if price is same as midprice. In that case, we use the following rule.

> The tick rule is the most commonly used level-1 algorithm. This rule is rather simple and classifies a trade as buyer-initiated if the trade price is above the preceding trade price (an uptick trade) and as seller-initiated if the trade price is below the preceding trade price (a downtick trade). If the trade price is the same as the previous trade price (a zero-tick trade), the rule looks for the closest prior price that differs from the current trade price. Zero-uptick trades are classified as buys, and zero-downtick trades are classified as sells.

The rule above can be simply implemented through the code below.

<div class = "pycode">
```{python}
ss = ss.assign(spread = ss['ask1'] - ss['bid1'],
               mid = (ss['bid1'] + ss['ask1'])*.5,
               ind = 0,
               price1 = ss['price'].shift(1, fill_value=0),
               price2 = ss['price'].shift(2, fill_value=0)
              )
              
buy = ((ss['price'] > ss['mid']) |
        ((ss['price'] == ss['mid']) & (ss['price'] > ss['price1'])) |
        ((ss['price'] == ss['mid']) & (ss['price'] == ss['price1']) & (ss['price'] > ss['price2'])))
        
ss.loc[buy, 'ind'] = 1
ss.loc[~buy, 'ind'] = -1

ss1 = ss[['time', 'price','size','ask1','bid1','ind','spread']]

ss1.head(5)
```
</div>


### We will discuss what to do more with using trading indicator value along with other variables extracted from tick data.

## Source

- 데이터 provided by 00 Team
- http://dee.uib.es/digitalAssets/234/234006_Pascual.pdf