---
title: Which one has higher volatility?
author: Roh
date: '2021-01-15'
slug: which-one-has-higher-volatility
categories:
  - finance
  - investment
tags: []
output:
  blogdown::html_page:
    toc: true
    toc_depth: 2
---
<style>
div.rcode pre { background-color:#fffff0; }
div.rcode pre.r { background-color:#EDF6FA; }
</style>

<style>
div.pycode pre { background-color:#fffff0; }
div.pycode pre.python { background-color:#FFE9E9; }

</style>
<style type="text/css">

body, td {
   font-size: 20px;
}
code.r{
  font-size: 12px;
}
pre {
  font-size: 18px
}
</style>

*__Note__: 이 페이지는 Volatility에 관한 내용입니다. 오류 및 문의사항은 metrics@kakao.com 으로 메일주시면 감사하겠습니다*

**데이터 자체에 대한 질문과 데이터 제공에 관한 문의는 000 으로 메일 주시면 감사하겠습니다**

**R code 블럭과 Python code 블럭은 다음과 같이 색깔로 구분하겠습니다. 결과창은 동일하게 Ivory 색 블럭으로 표시됩니다.**

<div class = "rcode">
```{r}
# "이것은 R 코드 입니다."
```
</div>

<div class = "pycode">
```{python}
# "이것은 Python 코드 입니다."
```
</div>

Think about $ X\times Y \in [5,10] \times [5,10] $ where X represents evenly spaced time (e.g., every one second) and Y represents the price of certain financial product. 

We often hear about the shapes appearing in the prices chart of financial product. 

We are interested in  the volatility of price series using the most commonly used measure that is variance. 

[5,10] 을 5분에서 10분 사이로 생각해보자. ㅈ

```{r}
x = seq(5, 10, length.out = 100)
x
```



