---
title: sampling method in detail
author: Roh
date: '2020-10-21'
slug: sampling-method-in-detail
categories:
  - sampling technique
tags:
  - mcmc
  - gibbs
output:
  blogdown::html_page:
    toc: true
    toc_depth: 6
# draft: true
---

This page talks about how to sample from the distribution. We will go over well known sampling techniques and try to implement with various examples. 

## Basic

What it means to calcaulte the variance with mean 0?

$$ Var[X]= \int_{-\infty}^{\infty} x^2 f(x)dx $$
when $f(x)$ is $N(1,2)$. Then $f(x)$ is  $$\frac{1}{\sqrt{2\pi\sigma^2}}\cdot\exp\left(-(x-\mu)^2/2\sigma^2\right)$$ where $\mu$ is 0 and $\sigma$ is 2. 

$$ Var[X]= \int_{-\infty}^{\infty} x^2 \frac{f(x)}{g(x)}g(x)dx $$
where $g(x)$ is N(0,3)+1 

- Integrand of $x$ along with density being integrated against $f(x)$

We are going to show the true mean, mean using simple monte carlo, and mean using importance sampling.

```{r}
set.seed(1234)
mu = 0
sigma = 2
X = rnorm(1e5, mu, sigma)
X2 = X^2
mean(X2); var(X2)
```

```{r}
exp(-1000)/(exp(-1000)+exp(-1001))
1/(1+exp(-1))
```


## Inverse Transform Sampling

If we know the exact form of the distribution, we can sample the points using the inverse CDF technique.

## Rejection Sampling

## Importance Sampling

- Not for generating samples. It is a method to estimate the expected value of a function directly

## 