---
title: 키움 API Tick Data - Task1 - Complete
author: Roh
date: '2021-01-20'
slug: api-tick-data-task1-day5
categories:
  - finance
  - Python
  - highfrequency
tags: []
output:
  blogdown::html_page:
    toc: true
    toc_depth: 4
---
<style>
div.rcode pre { background-color:#fffff0; }
div.rcode pre.r { background-color:#EDF6FA; }
</style>

<style>
div.pycode pre { background-color:#fffff0; }
div.pycode pre.python { background-color:#FFE9E9; }
</style>

<style type="text/css">

body, td {
   font-size: 20px;
}
code.r{
  font-size: 12px;
}
pre {
  font-size: 18px
}
</style>

*__Note__: 이 페이지는 키움 API를 통해 Real Time Tick Data를 받아온 후 데이터를 클리닝 하는 과정을 보여주고 있습니다. 오류 및 문의사항은 metrics@kakao.com 으로 메일주시면 감사하겠습니다*

**데이터 자체에 대한 질문과 데이터 제공에 관한 문의는 000 으로 메일 주시면 감사하겠습니다**

**R code 블럭과 Python code 블럭은 다음과 같이 색깔로 구분하겠습니다. 결과창은 동일하게 Ivory 색 블럭으로 표시됩니다.**

<div class = "rcode">
```{r}
# "이것은 R 코드 입니다."
```
</div>

<div class = "pycode">
```{python}
# "이것은 Python 코드 입니다."
```
</div>

* Package
```{r, warning=FALSE, message=FALSE, error=FALSE}
library(dplyr)
library(tidyverse)
library(DT)
library(reticulate) # Python
#py_install(packages = "matplotlib")
#py_install(packages = "pandas")
#py_install(packages = 'dfply')

options(scipen=999)
options(max.print = 99999999)
options(digits=10)
```

* Import Data

- Check NA

<div class = "rcode">
```{r, warning=FALSE, message=FALSE, error=FALSE}
wd = "G:/공유 드라이브/Project_TBD/Stock_Data/real_time/kiwoom_stocks/2021-01-14"
wd2 = "/Volumes/GoogleDrive/공유 드라이브/Project_TBD/Stock_Data/real_time/kiwoom_stocks/2021-01-20"
fn = list.files(path = wd2,
                pattern = '.*stocks_trade.*\\.csv') # 마지막 30분 틱데이터 가지고오기
path = paste(wd2,fn[1:4],sep = '/')

# data = readr::read_csv(file = path, 
#                        col_names = c('code','trade_date','timestamp','price','open','high','low',
#                                      'size','cum_size','ask1','bid1'))

tbl =
  list.files(path = wd2, pattern = '.*stocks_trade.*\\.csv')[2:8] %>%
  map_df(~readr::read_csv(paste(wd2,.,sep = '/'),
                          col_names = c('code','trade_date','timestamp','price','open','high','low','size','cum_size','ask1','bid1','rotation','bs_ration', 'mkt_type', 'mkt_cap'),
                          col_types = cols(.default="d", code = "c")
                   )
         )

sum(is.na(tbl))
```
</div>

<div class = "pycode">
```{python}
import numpy as np
import matplotlib.pyplot as plt
import pandas as pd
from datetime import datetime, timedelta
#py_install(packages = "matplotlib")
#py_install(packages = "pandas")
pd.options.display.float_format = '{:.4f}'.format
pd.set_option('display.max_rows', 100)
df_py = r.tbl
df_py.head()
```
</div>

```{r, include=FALSE, warning=FALSE, message=FALSE, error=FALSE}
rm(list = "tbl")
gc()
```

# Task1: 레이턴시 체크 및 Buy/Sell 구분하기 

## Samsung Stock

- KOSPI Code for Samsung Electronics is '005930'

<div class = "pycode">
```{python}
ss = df_py[df_py.code == '005930'].reset_index(drop=True)
ss.shape
```
</div>

### 1. Use time object in python

* Delete rows where "second" does not lie between 0 and 59.

* Substring the timstamp to get each time component.

<div class = "pycode">
```{python}
ss = ss[ss['timestamp'].apply(lambda x: int(str(int(x*1000000))[12:14])) < 60 ]

ss = ss.assign( microsecond = ss['timestamp'].apply(lambda x: str(int(x*1000000))[14:20]),
                second = ss['timestamp'].apply(lambda x: str(int(x*1000000))[12:14]),
                minute = ss['timestamp'].apply(lambda x: str(int(x*1000000))[10:12]),
                hour = ss['timestamp'].apply(lambda x: str(int(x*1000000))[8:10]),
                day = ss['timestamp'].apply(lambda x: str(int(x*1000000))[6:8]),
                month = ss['timestamp'].apply(lambda x: str(int(x*1000000))[4:6]),
                year = ss['timestamp'].apply(lambda x: str(int(x*1000000))[0:4]),
                kw_time = ss['trade_date'].apply(lambda x: str(int(x))))
ss.head()
```
</div>

* Transform data type to str to use substring to make the format like "%Y-%m-%d %H:%M:%S.%f"
* We calculate the latency by subtracting 'trade_date' from the 'timestamp'

<div class = "pycode">
```{python}
ss['time'] = ss.apply(lambda x: datetime(
                  year = int(x['year']),
                  month = int(x['month']),
                  day = int(x['day']),
                  hour = int(x['hour']),
                  minute = int(x['minute']),
                  second = int(x['second']),
                  microsecond = int(x['microsecond'])), #.strftime("%H:%M:%S.%f")[:-3], 
                  axis=1)
ss.head(10)
```
</div>

### 2. Calculate latency and clean it [레이턴시 계산 및 정리]

* Calculate the latency by subtracting the time provided by Kiwoom dataset and self-recorded time at the moment of processing the data delivered through the Kiwoom API.

#### 오전 8시 9시 앞에 '0'을 붙여주기

<div class = "pycode">
```{python}
ss['kw_time'].apply(lambda x: x[0:2]).value_counts()
```
</div>


<div class = "pycode">
```{python}
ss['kw_time'] = ss.kw_time.apply(lambda x: '0' + x if ((x[0]=='8') | (x[0]=='9')) else x)
ss['kw_time'].apply(lambda x: x[0:2]).value_counts()
```
</div>

* 키움 타임 만들기

<div class = "pycode">
```{python}
ss['kwtime'] = ss.apply(lambda x: datetime(year = int(x['year']),
                                           month = int(x['month']),
                                           day = int(x['day']),
                                           hour = int(x['kw_time'][0:2]),
                                           minute = int(x['kw_time'][2:4]),
                                           second = int(x['kw_time'][4:6])),
                                           axis = 1
                                           )
```
</div>

#### Latency +/- 정보 가지고 오기

* + 인 경우, 로컬 타임이 키움 타임보다 늦는 경우
* - 인 경우, 로컬 타임이 키움 타임보다 빠른 경우 (weird)  

<div class = "pycode">
```{python}
ss['latency_ind']=ss.apply(lambda x: 1 if (int(str(int(x['timestamp']*1000000))[8:14]) > x['trade_date']) else
                            0 if (int(str(int(x['timestamp']*1000000))[8:14]) == x['trade_date']) else -1, axis=1)
```
</div>


<div class = "pycode">
```{python}
plt.scatter(ss['latency_ind'].reset_index().index,ss['latency_ind'])
plt.title('latency +/-')
```
</div>

#### Latency Value 가지고 오기

<div class = "pycode">
```{python}
ss['latency_val'] = ss.apply(lambda x: abs(x['time'].second - x['kwtime'].second),
                        axis =1
                        )
ss['latency'] = ss['latency_val']*ss['latency_ind']
ss1 = ss[['time', 'kwtime', 'kw_time', 'hour', 'minute', 'second', 'price','size','ask1','bid1','latency']]
ss1.head(10)
```
</div>

<div class = "pycode">
```{python}
plt.scatter(ss1.reset_index().index, ss1.latency)
plt.title('Latency Over Time Before Cleaning')
```
</div>

<div class = "pycode">
```{python}
ss1.latency.value_counts().head(10).iloc[::-1].plot.barh(stacked=True)
plt.title('Latency Top 10 Bar Plot')
```
</div>

<div class = "pycode">
```{python}
ss1.latency.value_counts().index
```
</div>

* Check the time where large discrepancy happened!

#### over 30 or not!

- We assume that when the latency is minus value, the maximum time difference is 1.
- 음의 레이턴시가 나오는 경우는 최악의 경우 1초 빠르게 record 하는 경우!!
- When latency is a number x (e.g. 59) over 30, the actual latency is 60 - x
- When latency is a number x (e.g. 20) below 30 and over 0, the actual latency is x
- When latency is a number x (e.g. -59) below (let's say) -5, the actual latency is -(60 + x)
- We can check it as in below.


<div class = "pycode">
```{python}
ss1[ss1.latency == -59][['kwtime','hour','minute','second','latency']]
ss1[ss1.latency == 40][['kwtime','hour','minute','second','latency']]
ss1[ss1.latency == 39][['kwtime','hour','minute','second','latency']]
ss1[ss1.latency == 21][['kwtime','hour','minute','second','latency']]
ss1[ss1.latency == 20][['kwtime','hour','minute','second','latency']]
```
</div>

#### Correct the wrongly calculated latency 

<div class = "pycode">
```{python}
ss2 = ss1.copy()
ss2['latency']=ss1.apply(lambda x: 60-x['latency'] if (x['latency'] > 30) else \
                                    (-(60+x['latency']) if (x['latency']<-5) else x['latency']),axis=1)
```
</div>

<div class = "pycode">
```{python}
# x coordinates for the lines

xcoords = [ ss2.index[ss2['kw_time'] == '090500'][0], 
            ss2.index[ss2['kw_time'] == '091000'][0],
            ss2.index[ss2['kw_time'] == '092000'][0] ]

# colors for the lines
colors = ['darkgreen','green','limegreen']

time = ['09:05:00','09:10:00','09:20:00']

plt.scatter(ss2.reset_index().index, ss2.latency)
for xc,c,time in zip(xcoords,colors,time):
    plt.axvline(x=xc, label='time = {}'.format(time), c=c)
plt.legend()
plt.title('Latency Over Time After Cleaning')
```
</div>

<div class = "pycode">
```{python}
ss2.latency.value_counts().iloc[::-1].plot.barh(stacked=True)
plt.title('Latency Top 10 Bar Plot')
```
</div>

***

## Hyundai Car Stock

- KOSPI Code for Samsung Electronics is '005380'

<div class = "pycode">
```{python}
ss = df_py[df_py.code == '005380'].reset_index(drop=True)
ss.shape
```
</div>

### 1. Use time object in python

* Delete rows where "second" does not lie between 0 and 59.

* Substring the timstamp to get each time component.

<div class = "pycode">
```{python}
ss = ss[ss['timestamp'].apply(lambda x: int(str(int(x*1000000))[12:14])) < 60 ]

ss = ss.assign( microsecond = ss['timestamp'].apply(lambda x: str(int(x*1000000))[14:20]),
                second = ss['timestamp'].apply(lambda x: str(int(x*1000000))[12:14]),
                minute = ss['timestamp'].apply(lambda x: str(int(x*1000000))[10:12]),
                hour = ss['timestamp'].apply(lambda x: str(int(x*1000000))[8:10]),
                day = ss['timestamp'].apply(lambda x: str(int(x*1000000))[6:8]),
                month = ss['timestamp'].apply(lambda x: str(int(x*1000000))[4:6]),
                year = ss['timestamp'].apply(lambda x: str(int(x*1000000))[0:4]),
                kw_time = ss['trade_date'].apply(lambda x: str(int(x))))
ss.head()
```
</div>

* Transform data type to str to use substring to make the format like "%Y-%m-%d %H:%M:%S.%f"
* We calculate the latency by subtracting 'trade_date' from the 'timestamp'

<div class = "pycode">
```{python}
ss['time'] = ss.apply(lambda x: datetime(
                  year = int(x['year']),
                  month = int(x['month']),
                  day = int(x['day']),
                  hour = int(x['hour']),
                  minute = int(x['minute']),
                  second = int(x['second']),
                  microsecond = int(x['microsecond'])), #.strftime("%H:%M:%S.%f")[:-3], 
                  axis=1)
ss.head(10)
```
</div>

### 2. Calculate latency and clean it [레이턴시 계산 및 정리]

* Calculate the latency by subtracting the time provided by Kiwoom dataset and self-recorded time at the moment of processing the data delivered through the Kiwoom API.

#### 오전 8시 9시 앞에 '0'을 붙여주기

<div class = "pycode">
```{python}
ss['kw_time'].apply(lambda x: x[0:2]).value_counts()
```
</div>


<div class = "pycode">
```{python}
ss['kw_time'] = ss.kw_time.apply(lambda x: '0' + x if ((x[0]=='8') | (x[0]=='9')) else x)
ss['kw_time'].apply(lambda x: x[0:2]).value_counts()
```
</div>

* 키움 타임 만들기

<div class = "pycode">
```{python}
ss['kwtime'] = ss.apply(lambda x: datetime(year = int(x['year']),
                                           month = int(x['month']),
                                           day = int(x['day']),
                                           hour = int(x['kw_time'][0:2]),
                                           minute = int(x['kw_time'][2:4]),
                                           second = int(x['kw_time'][4:6])),
                                           axis = 1
                                           )
```
</div>

#### Latency +/- 정보 가지고 오기

* + 인 경우, 로컬 타임이 키움 타임보다 늦는 경우
* - 인 경우, 로컬 타임이 키움 타임보다 빠른 경우 (weird)  

<div class = "pycode">
```{python}
ss['latency_ind']=ss.apply(lambda x: 1 if (int(str(int(x['timestamp']*1000000))[8:14]) > x['trade_date']) else -1,
        axis=1)
```
</div>


<div class = "pycode">
```{python}
plt.scatter(ss['latency_ind'].reset_index().index,ss['latency_ind'])
plt.title('latency +/-')
```
</div>

#### Latency Value 가지고 오기

<div class = "pycode">
```{python}
ss['latency_val'] = ss.apply(lambda x: abs(x['time'].second - x['kwtime'].second),
                        axis =1
                        )
ss['latency'] = ss['latency_val']*ss['latency_ind']
ss1 = ss[['time', 'kwtime', 'kw_time', 'hour', 'minute', 'second', 'price','size','ask1','bid1','latency']]
ss1.head(10)
```
</div>

<div class = "pycode">
```{python}
plt.scatter(ss1.reset_index().index, ss1.latency)
plt.title('Latency Over Time Before Cleaning')
```
</div>

<div class = "pycode">
```{python}
ss1.latency.value_counts().head(10).iloc[::-1].plot.barh(stacked=True)
plt.title('Latency Top 10 Bar Plot')
```
</div>

<div class = "pycode">
```{python}
ss1.latency.value_counts().index
```
</div>

* Check the time where large discrepancy happened!

#### over 30 or not!

- We assume that when the latency is minus value, the maximum time difference is 1.
- 음의 레이턴시가 나오는 경우는 최악의 경우 1초 빠르게 record 하는 경우!!
- When latency is a number x (e.g. 59) over 30, the actual latency is 60 - x
- When latency is a number x (e.g. 20) below 30 and over 0, the actual latency is x
- When latency is a number x (e.g. -59) below (let's say) -5, the actual latency is -(60 + x)
- We can check it as in below.


<div class = "pycode">
```{python}
ss1[ss1.latency == -59][['kwtime','hour','minute','second','latency']]
ss1[ss1.latency == 40][['kwtime','hour','minute','second','latency']]
ss1[ss1.latency == 39][['kwtime','hour','minute','second','latency']]
ss1[ss1.latency == 21][['kwtime','hour','minute','second','latency']]
ss1[ss1.latency == 20][['kwtime','hour','minute','second','latency']]
```
</div>

#### Correct the wrongly calculated latency 

<div class = "pycode">
```{python}
ss2 = ss1.copy()
ss2['latency']=ss1.apply(lambda x: 60-x['latency'] if (x['latency'] > 30) else \
                                    (-(60+x['latency']) if (x['latency']<-5) else x['latency']),axis=1)
```
</div>

<div class = "pycode">
```{python}
# x coordinates for the lines

xcoords = [ ss2.index[ss2['kw_time'] == '090500'][0], 
            ss2.index[ss2['kw_time'] == '091000'][0],
            ss2.index[ss2['kw_time'] == '092000'][0] ]

# colors for the lines
colors = ['darkgreen','green','limegreen']

time = ['09:05:00','09:10:00','09:20:00']

plt.scatter(ss2.reset_index().index, ss2.latency)
for xc,c,time in zip(xcoords,colors,time):
    plt.axvline(x=xc, label='time = {}'.format(time), c=c)
plt.legend()
plt.title('Latency Over Time After Cleaning')
```
</div>

<div class = "pycode">
```{python}
ss2.latency.value_counts().iloc[::-1].plot.barh(stacked=True)
plt.title('Latency Top 10 Bar Plot')
```
</div>

***

## Source

- 데이터 provided by 00 Team
---